%\begin{table}[h]
%\centering
%\caption{Test Case Descriptions}
%\label{my-label}
\begin{scriptsize}
\begin{longtable}{@{}p{0.04\textwidth}p{0.17\textwidth}p{0.17\textwidth}p{0.17\textwidth}p{0.14\textwidth}p{0.15\textwidth}@{}}
\caption{Test Case Descriptions}\\
\label{testcasetabel}\\
\toprule
ID & Objective & Precondition & Steps & Test Data & Expected \newline Result \\ \midrule
TC1 & Create new classifier tab & There is only one tab in the classifier tab box. & - Press the '+' button above the classifier tab box. & N/A & A new tab is created with the number 2 on it. \\
TC2 & Open dataset window. & N/A & - Press the 'Select Dataset..' button in the main window. & N/A & Dataset window is shown. \\ 
TC3 & Select training dataset. & Dataset window is open. & - Press the 'Select Training Set..' button. \newline - Select a file. & Any .csv file. & Absolute path of the file appears in the textbox below 'Training Dataset' label. \\ 
TC4 & Deselect k-Fold Cross Validation. & Dataset window is open and k-Fold Cross Validation checkbox is checked. & - Press the k-Fold Cross Validation checkbox. & N/A & Testing dataset label, textbox and button become enabled and folds spinbox is disabled. \\ 
TC5 & Select testing dataset. & Dataset window is open. & - Press the 'Select Testing Set...' button. \newline - Select a file. & Any valid .csv file. &  Absolute path of the file appears in the textbox below 'Training Dataset' label.\\ 
TC6 & Select dataset fields. & Dataset window is open.  &  - Press the 'Select Dataset Fields' button. \newline - Select a file.&  A .csv file containing two columns, the second of which is one of the values: Numeric, Nominal, Binary. &  Absolute path of the file appears in the textbox below 'Dataset Fields' label, and the 'Numeric', 'Nominal', and 'Binary' list views are populated.\\ 
TC7 & Select attack categories. & Dataset window is open. & - Press the 'Select Attack Categories' button. \newline - Select a file. &  Any valid .csv file.&  Absolute path of the file appears in the textbox below 'Attack Categories' label.\\ 
TC8 & Display dataset info in main window. & Valid dataset files have been selected and dataset window is open. & - Press the OK button in the bottom right of the dataset window. & N/A & The dataset window closes, showing no error message and the textbox in the top left of the main window is populated with information regarding the dataset. \\ 
TC9 & Select classifier. & N/A & - Press the 'Select Classifier' button below the 'Classifier One' label. \newline - Select a file. & Any .py file. & Absolute path of the file appears in the textbox below the 'Classifier One' label. \\ 
TC10 & Select two stage classification. & Two stage classification checkbox is unchecked. & - Press the 'Two Stage Classification' checkbox. & N/A & The 'Second Classifier' label, textbox, and 'Select Classifier' button are enabled. \\ 
TC11 & Select second classifier. & The 'Two Stage Classification' checkbox is checked. & - Press the 'Select Classifier' button below the 'Classifier Two' label. \newline - Select a file. & Any .py file. & Absolute path of the file appears in the textbox below the 'Classifier One' label. \\ 
TC12 & Error message on invalid training dataset filename. & N/A & - Enter a filepath which does not exist into the textbox below the 'Training Dataset' label. \newline - Supply every other textbox with valid filepaths. \newline - Press the OK button in the bottom right of the dataset window. & N/A & An error message is shown and the window does not exit. \\ 
TC13 & Error message on invalid testing dataset filename. & N/A & - Enter a filepath which does not exist into the textbox below the 'Testing Dataset' label. \newline - Supply every other textbox with valid filepaths. \newline - Press the OK button in the bottom right of the dataset window. & N/A & An error message is shown and the window does not exit. \\ 
TC14 & Error message on invalid field names filename. & N/A & - Enter a filepath which does not exist into the textbox below the 'Dataset Fields' label. \newline - Supply every other textbox with valid filepaths. \newline - Press the OK button in the bottom right of the dataset window. & N/A & An error message is shown and the window does not exit. \\ 
TC15 & Error message on invalid attack categories filename. & N/A & - Enter a filepath which does not exist into the textbox below the 'Attack Categories' label. \newline - Supply every other textbox with valid filepaths. \newline - Press the OK button in the bottom right of the dataset window. & N/A & An error message is shown and the window does not exit. \\ 
TC16 & Error message on invalid training dataset. & Valid dataset files have been selected, except from training dataset which is invalid. A valid classifier configuration has been selected. & - Press the 'Run' button on the main window. & N/A & An error message is displayed and the program does not crash. \\ 
TC17 & Error message on invalid testing dataset. & Valid dataset files have been selected, except from testing dataset which is invalid. A valid classifier configuration has been selected. & - Press the 'Run' button on the main window. & N/A & An error message is displayed and the program does not crash. \\ 
TC18 & Error message on invalid field names. & Valid dataset files have been selected, except from dataset fields which do not correspond to the dataset. A valid classifier configuration has been selected. & - Press the 'Run' button on the main window. & N/A & An error message is displayed and the program does not crash. \\ 
TC19 & Error message on invalid attack categories. & Valid dataset files have been selected, except from attack categories which do not correspond to the dataset. A valid classifier configuration has been selected. & - Press the 'Run' button on the main window. & N/A & An error message is displayed and the program does not crash. \\ 
TC20 & Dataset window is populated with existing information. & Dataset file paths have been previously selected and the OK button pressed on the dataset window. & - Press the 'Select Dataset..' button on the main window. & N/A & The text boxes containing the paths to files should all be filled with the information which was entered previously, as well as fold count and stochastic classifier run count. \\ 
TC21 & Error message on invalid classifier. & N/A & - Press the 'Select Classifier' button the main window. \newline - Select an invalid .py file. \newline - Press the 'Run' button. & An invalid .py file. & An error message is displayed and the program does not crash. \\ 
TC22 & Delete classifier tab. & There is more than one tab in the classifier tab. & - Press the 'x' button next to the tab number in the classifiers tabs in any tab except the first. & N/A & The tab in which the 'x' button was pressed should close. \\ 
TC23 & Show results for all classifiers run. & Valid dataset has been selected and valid classifier configuration has been selected. & - Press the 'Run' button. & N/A & A tab should be created on the main tab container for each classifier configuration run, and populated with classification information. \\ 
TC24 & Show graph of classification results. & Classifiers have been run and results retrieved. & - Select the 'Graphs' tab on the main window. \newline - Select a class from the drop down in the 'Graphs' tab. \newline - Press the 'Show' button. & N/A & A graph is shown which has a series for each classifier configuration specified. \\ 

 \bottomrule
\end{longtable}
\end{scriptsize}
%\end{table}
